---
title: "Join Lab"
author: "Jonathan Aguilar"
date: "DATA 2401"
output: html_document
---


# Joins and `dplyr` practice

Load the `"nycflights13"` package.  You'll also need to load `dplyr` or `tidyverse`, if you didn't above.

```{r}
library(tidyverse)
install.packages("nycflights13")
library(nycflights13)
my_flights <- flights # store my own version
my_flights <- na.omit(my_flights) ## this dataset has data on cancelled flights too, 
                                  ## so we 'll remove those for now

```


Create a dataframe of the average arrival delays for each _destination_

```{r}
arr_delay <- my_flights %>% group_by(dest) %>% summarise(avg_delay = mean(arr_delay ))
arr_delay
```

Then use `left_join()` to join on the "airports" dataframe, which has the airport information.

```{r}
arr_delay <- left_join(arr_delay, airports, join_by(dest == faa))
arr_delay
```


Note here: the column names don't match anymore! 

Which airport had the largest average arrival delay?

```{r}
largest_avg_arrival_delay <- arr_delay %>% summarise(Airport = arr_delay$name.x, max_delay = max(avg_delay))

largest_avg_arrival_delay
#Columbia Metropolitan
```

What is the average departure delay for each airline? To answer this with something more than just the FAA code of the airline name, you need to create a dataframe of the average departure delay for each _airline_, then use `left_join()` to join on the "airlines" dataframe


```{r}
avg_depart_delay <- my_flights %>% group_by(carrier) %>% summarise(avg_depart_delay = mean(dep_delay))
avg_depart_delay
```

Which airline had the smallest average departure delay? 



```{r}
smallest_departure <- left_join(avg_depart_delay, airlines, by = "carrier")
smallest_departure

smallest_avg_depart_airport <- smallest_departure %>% slice_min(avg_depart_delay)
smallest_avg_depart_airport
```





Examples from R4DS:    

```{r}

top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest

```

Now you want to find each flight that went to one of those destinations. You could, of course, filter. But! Now you can semi-join: 

```{r}
flights %>% 
  semi_join(top_dest)
```


What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?
Perform the two joins below and consider what the results are. 
```{r}
anti_join(flights, airports, by = c("dest" = "faa"))
#returns all information on flights where the destination does not match the faa in airports
```
```{r}
anti_join(airports, flights, by = c("faa" = "dest"))
#returns all information on airlines where faa does not match the destination on flights
```